### 2016-03-16 misc/notes.tex
implement SWAP_NVST using one 'cheating' operation

We introduce a new notation for operations:
\begin{align}
  U\sp{t_0H}=\exp\p{-it_0 H}.
\end{align}
In this notation,
\begin{align}
  R_{\uv n}\p\phi=U\sp{\f\phi2\sigma_{\uv n}},
\end{align}
\begin{align}
  \t{cNOT}^{m,n}=U\sp{\f\pi4\p{\sigma_z^m+\sigma_x^n
      -\sigma_z^m\sigma_x^n}}
  =U\sp{\f\pi4\sigma_z^m}U\sp{\f\pi4\sigma_x^n}
  U\sp{-\f\pi4\sigma_z^m\sigma_x^n}
\end{align}
\begin{align}
  i\t{SWAP}^{m,n}=U\sp{-\f\pi4\p{\sigma_x^m\sigma_x^n
      +\sigma_y^m\sigma_y^n}} =U\sp{-\f\pi4\sigma_y^m\sigma_y^n}
  U\sp{-\f\pi4\sigma_x^m\sigma_x^n},
\end{align}
\begin{align}
  \E^{m,n}=U\sp{-\f\pi4\sigma_y^m\sigma_x^n},
\end{align}
and therefore
\begin{multline}
  \t{SWAP}^{NV,ST}= U\sp{\f\pi4\sigma_z^{NV}}U\sp{\f\pi4\sigma_z^1}
  U\sp{\f\pi4\sigma_x^{NV}\sigma_x^1}
  U\sp{\f\pi4\sigma_y^{NV}\sigma_y^1}
  U\sp{-\f\pi4\sigma_y^{NV}\sigma_x^2} \\
  U\sp{\f\pi4\sigma_z^{NV}}U\sp{\f\pi4\sigma_x^1}
  U\sp{-\f\pi4\sigma_z^{NV}\sigma_x^1}
  U\sp{\f\pi4\sigma_y^{NV}\sigma_x^2} \\
  U\sp{-\f\pi4\sigma_y^{NV}\sigma_y^1}
  U\sp{-\f\pi4\sigma_x^{NV}\sigma_x^1}
  U\sp{-\f\pi4\sigma_z^1}U\sp{-\f\pi4\sigma_z^{NV}}.
\end{multline}

@@

Individual nuclear control is a slow operation, whereas operations of
the form $U\sp{\phi\sum_j\sigma_{\uv n}^j}$ can be performed
relatively quickly and with high fidelity. We therefore modify the
above $\t{SWAP}^{NV,ST}$ decomposition as
\begin{multline}
  \t{SWAP}^{NV,ST}=
  U\sp{\f\pi4\sigma_z^{NV}}U\sp{\f\pi4\sum_j\sigma_z^j}
  U\sp{\f\pi4\sigma_x^{NV}\sigma_x^1}
  U\sp{\f\pi4\sigma_y^{NV}\sigma_y^1}
  U\sp{\f\pi4\sigma_y^{NV}\sigma_y^2} \\
  U\sp{\f\pi4\sigma_z^{NV}}U\sp{\f\pi4\sigma_x^1}
  U\sp{-\f\pi4\sigma_z^{NV}\sigma_x^1}
  U\sp{-\f\pi4\sigma_y^{NV}\sigma_y^2} \\
  U\sp{-\f\pi4\sigma_y^{NV}\sigma_y^1}
  U\sp{-\f\pi4\sigma_x^{NV}\sigma_x^1}
  U\sp{-\f\pi4\sum_j\sigma_z^j}U\sp{-\f\pi4\sigma_z^{NV}}.
\end{multline}

@@

\section*{Coherence of a single qubit}

Any single-qubit density matrix $\rho$ can be decomposed as
\begin{align}
  \rho=\f12\p{I+\v v\c\v\sigma}=\f12\m{1+v_z&v_x-iv_y\\v_x+iv_y&1-v_z}.
\end{align}
for a vector $\v v$ with $\v v\c\v v\le 1$. The components $v^i$ of
$\v v$ can be recovered via the anticommutator
\begin{align}
  v^i=\f12\set{\sigma_i,2\rho-I}.
\end{align}
Question: is $C_p\p\rho=\abs{\v v}$ a proper measure of
coherence?

\subsection*{$l_1$ norm of coherence}

In general,
\begin{align}
  C_{l_1}\p\rho=\sum_{i\ne j}\abs{\rho_{ij}}
  =\f12\p{\abs{v_x-iv_y}+\abs{v_x+iv_y}}=\sqrt{v_x^2+v_y^2}.
\end{align}
Equivalently,
\begin{align}
  C_{l_1}\p\rho=\abs{\v v-\v v\c\v z\uv z}
  =\abs{P\sp{\uv v,\uv x\w\uv y}}=v\sin\theta,
\end{align}
where $\theta\in\sp{0,\pi}$ is the polar angle of $\v v$. For any
given state, therefore, this measure of coherence is thus minimal when
when $\v v\parallel\uv z$ (i.e. $\theta\in\set{0,\pi}$, so
$C_{l_1}\p\rho=0$), and maximal when $\v v\perp\uv z$
(i.e. $\theta=\pi/2$, so $C_{l_1}\p\rho=v$).

It is worthwhile to notice that when $\v v\perp\uv z$,
$C_{l_1}\p\rho=C_p\p\rho$, but that $C_p\p\rho$ is
basis-independent.

\subsection*{Relative entropy of coherence}

The relative entropy of coherence is
\begin{align}
  C_{\t{rel. ent.}}\p\rho=S\p{\rho_{\t{diag}}}-S\p\rho,
\end{align}
The von Neumann entropy is
\begin{align}
  S\p\rho=-\tr\p{\rho\log\rho}.
\end{align}
If $\rho=\sum_j\eta_j\op j$, then $S\p\rho=-\sum_j\eta_j\log\eta_j$. We
can express
\begin{align}
  \rho_{\t{diag}}=\f12\p{I+v^z\sigma_z}
  =\f12\sp{\p{1+v^z}\op{+_z}+\p{1-v^z}\op{-_z}},
\end{align}
and
\begin{align}
  \rho=\f12\p{I+v\sigma_v}
  =\f12\sp{\p{1+v}\op{+_v}+\p{1-v}\op{-_v}}.
\end{align}
Letting $v_\pm\equiv \p{1\pm v}/2$ and $v^z_\pm\equiv\p{1\pm
  v^z}/2=\p{1\pm v\cos\theta}/2$ (where $\theta$ is the polar angle of
$\v v$),
\begin{align}
  \rho_{\t{diag}}=v^z_+\op{+_z}+v^z_-\op{-_z}, &&
  \rho=v_+\op{+_v}+v_-\op{-_v},
\end{align}
which means
\begin{align}
  C_{\t{rel. ent.}}\p\rho
  =-\p{v^z_+\log v^z_++v^z_-\log v^z_-}+\p{v_+\log v_++v_-\log v_-}.
\end{align}
For any given state, this measure of coherence is similarly minimal
when $\v v\parallel\uv z$ (i.e. $\rho_{\t{diag}}=\rho$, so
$C_{\t{rel. ent.}}=0$), and maximal when $\v v\perp\uv z$ (i.e.
$v^z=0$ and $v^z_\pm=1/2$, so $C_{\t{rel. ent.}}\p\rho=\log2-S\p\rho$).

\section*{Coherence of many qubits}

Consider a density matrix of the form
\begin{align}
  \rho=\f1{2^N}\p{I
    +\sum_{\substack{i_0,i_1,\cdots,i_{N-1}\\\exists n:i_n\ne0}}
    v^{i_0i_1\cdots i_{N-1}}\bigotimes_{n<N}\sigma_{i_n}},
\end{align}
which identifies the state of $N$ qubits. Here all
$i_n\in\set{0,1,2,3}$,
$\sigma_{i_n}\in\set{I,\sigma_x,\sigma_y,\sigma_z}$, and $n$ indexes
the qubit upon which $\sigma_{i_n}$ acts.

Questions: What restrictions exist on $\rho$? What do these
restrictions imply about $\v v$?

>>>>

We introduce a new notation for operations:
\begin{align}
  U\sp{t_0H}=\exp\p{-it_0 H}.
\end{align}
In this notation,
\begin{align}
  R_{\uv n}\p\phi=U\sp{\f\phi2\sigma_{\uv n}},
\end{align}
\begin{align}
  \t{cNOT}^{m,n}=U\sp{\f\pi4\p{\sigma_z^m+\sigma_x^n
      -\sigma_z^m\sigma_x^n}} =R_z^m\p{\f\pi2} R_x^1\p{\f\pi2}
  U\sp{-\f\pi4\sigma_z^m\sigma_x^n}
\end{align}
\begin{align}
  i\t{SWAP}^{m,n}=U\sp{-\f\pi4\p{\sigma_x^m\sigma_x^n
      +\sigma_y^m\sigma_y^n}} =U\sp{-\f\pi4\sigma_y^m\sigma_y^n}
  U\sp{-\f\pi4\sigma_x^m\sigma_x^n},
\end{align}
\begin{align}
  \E^{m,n}=U\sp{-\f\pi4\sigma_y^m\sigma_x^n},
\end{align}
and therefore, defining
\begin{align}
  \mathcal M &={\E^\dag}^{NV,2} i\t{SWAP}^{NV,1}
  R_z^1\p{-\f\pi2}R_z^{NV}\p{-\f\pi2} \tag*{} \\
  &=U\sp{\f\pi4\sigma_y^{NV}\sigma_x^2}
  U\sp{-\f\pi4\sigma_y^{NV}\sigma_y^1}
  U\sp{-\f\pi4\sigma_x^{NV}\sigma_x^1}
  R_z^1\p{-\f\pi2}R_z^{NV}\p{-\f\pi2}
\end{align}
we can express
\begin{align}
  \t{SWAP}^{NV,ST}=\mathcal M^\dag R_z^{NV}\p{\f\pi2} R_x^1\p{\f\pi2}
  U\sp{-\f\pi4\sigma_z^{NV}\sigma_x^1} \mathcal M.
\end{align}

@@

When addressing larmor pairs with frequency $\omega_\l$, operations of
the form $\prod_{j\in\set{k:\omega_k=\omega_\l}}R_{\uv n}^j\p\phi$ are
easier to realize than individual rotations $R_{\uv n}^m\p\phi$. We
therefore define
\begin{align}
  \mathcal M'=U\sp{-\f\pi4\sigma_y^{NV}\sigma_y^2}
  U\sp{-\f\pi4\sigma_y^{NV}\sigma_y^1}
  U\sp{-\f\pi4\sigma_x^{NV}\sigma_x^1}
  \sp{\prod_{j\in\set{k:\omega_k=\omega_\l}}R_z^j\p{-\f\pi2}}
  R_z^{NV}\p{-\f\pi2}
\end{align}
and express $\t{SWAP}^{NV,ST}$ as
\begin{align}
  \t{SWAP}^{NV,ST}={\mathcal M'}^\dag R_z^{NV}\p{\f\pi2} R_x^1\p{\f\pi2}
  U\sp{-\f\pi4\sigma_z^{NV}\sigma_x^1} \mathcal M'.
\end{align}

@@

\section*{Addressing larmor pairs}

When applying a static control field $\v B_{\t{ctl}}$, we have the
effective interaction + control Hamiltonian
\begin{align}
  H=\f12m_s\sigma_z\sum_j\v A_j\c\v I_j-\sum_j\gamma_j\v
  B_{\t{ctl}}\c\v I_j
\end{align}
